---
title: "Game Dataset EDA"
author: "Tamojit Roy"
date: "22/11/2021"
output: html_document
---

## Reading the game dataset

```{r}
df = read.csv(file="Datasets/clean_game.csv")
```

## Preprocessing the all reviews column

```{r}
#install.packages("micropan")
library(micropan)

df1 <- df[-which(df$all_reviews == ""), ]

n = count(df1)

review = c()
for(i in (1:28470)) {
  str=df$all_reviews[i]
  str2=unlist(gregexpr(pattern='-',str))
  vex=substr(str,str2+2,str2+3)
  review <- append(review,strtoi(vex))
}

review <- data.frame(review)
review <- review[-which(is.na(review$review)), ]
```

## Distribution of Game Reviews

```{r}
#install.packages("ggplot2")
library(ggplot2)
ggplot() + aes(review)+ geom_histogram(binwidth=1, colour="blue", fill="blue")

# ggplot(review,aes(x=V1))+geom_histogram()+labs(title="Distribution of Game Reviews",y="Number of Games",x="Rating %")
```

## Preprocessing of genre column

```{r}
df2 <- df[-which(df$genre == ""), ]

n = count(df2)

genre1 <- df2$genre

library(base)

list1 <- unlist(strsplit(genre1[1], ","))
genres = c()

for(i in (2:40395)) {
  list2 <- unlist(strsplit(genre1[i], ","))
  genres = union(list1,list2)
  list1 = genres
}

no_of_games = replicate(34,0)

genre2 <- data.frame(genres,no_of_games)
```

## 

```{r}
library(stringr)

for(i in (1:40395)) {
  list_store <- unlist(strsplit(genre1[i], ","))
  for(j in (1:34)) {
    if(sum(str_detect(list_store,genres[j])) == 1) {
      genre2$no_of_games[j] = genre2$no_of_games[j]+1
    }
  }
}
```

## All distinct genres

```{r}
genres
```

## Plotting recurrence of genre bar chart

```{r}
ggplot(data=genre2, aes(x=no_of_games, y=genres)) + geom_bar(stat="identity") + ggtitle("Recurrence of Genre")
```

## Preprocessing popular tags column

```{r}
df3 <- df[-which(df$popular_tags == ""), ]

n = count(df3)

tags1 <- df3$popular_tags

library(base)

list1 <- unlist(strsplit(tags1[1], ","))
popular_tags = c()

for(i in (2:37888)) {
  list2 <- unlist(strsplit(tags1[i], ","))
  popular_tags = union(list1,list2)
  list1 = popular_tags
}

no_of_games = replicate(376,0)

tags2 <- data.frame(popular_tags,no_of_games)
```

## Distinct Popular tags

```{r}
popular_tags
```

## 

```{r}
library(stringr)

for(i in (1:37888)) {
  list_store <- unlist(strsplit(tags1[i], ","))
  for(j in (1:376)) {
    if(sum(str_detect(list_store,popular_tags[j])) == 1) {
      tags2$no_of_games[j] = tags2$no_of_games[j]+1
    }
  }
}
```

## Plotting recurrence of popular tags bar chart

```{r}
arrange(tags2,desc(no_of_games))
tags3 = head(tags2,20)

ggplot(data=tags3, aes(x=no_of_games, y=popular_tags)) + geom_bar(stat="identity") + ggtitle("Recurrence of Popular Tags")
```

## Preprocessing game details column

```{r}
df4 <- df[-which(df$game_details == ""), ]

n = count(df4)

details1 <- df4$game_details

library(base)

list1 <- unlist(strsplit(details1[1], ","))
game_details = c()

for(i in (2:40313)) {
  list2 <- unlist(strsplit(details1[i], ","))
  game_details = union(list1,list2)
  list1 = game_details
}

no_of_games = replicate(34,0)

details2 <- data.frame(game_details,no_of_games)
```

## Distinct game details

```{r}
game_details
```

## 

```{r}
library(stringr)

for(i in (1:40313)) {
  list_store <- unlist(strsplit(details1[i], ","))
  for(j in (1:34)) {
    if(sum(str_detect(list_store,game_details[j])) == 1) {
      details2$no_of_games[j] = details2$no_of_games[j]+1
    }
  }
}
```

## Plotting recurrence of game details bar chart

```{r}
arrange(details2,desc(no_of_games))
details3 = head(details2,20)

ggplot(data=details3, aes(x=no_of_games, y=game_details)) + geom_bar(stat="identity") + ggtitle("Recurrence of Game Details")
```

## 

```{r}

```

## 

```{r}

```

## 

```{r}

```
